### 2024-01-27 manuscript/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=2.54cm,right=2.54cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\usepackage{lipsum} \usepackage{libertine}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Urban mobility in Latin America after COVID-19},
  pdfauthor={Carmen Cabrera-Arnau; Francisco Rowe; Miguel González-Leonardo; Andrea Nasuto; Ruth Neville},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{Urban mobility in Latin America after COVID-19}}

\author[1]{Carmen Cabrera-Arnau}
\author[1]{Francisco Rowe}
\author[2]{Miguel González-Leonardo}
\author[1]{Andrea Nasuto}
\author[1]{Ruth Neville}

\affil[1]{Geographic Data Science Lab, Department of Geography and
Planning, University of Liverpool, Liverpool, UK}
\affil[2]{Centre for Demographic Urban and Environmental Studies, El
Colegio de México, Ciudad de México, México}

\date{}
\begin{document}
\maketitle
\begin{abstract}
The movement of people between locations is crucial for sustainable and
inclusive cities, facilitating the exchange of knowledge and resources.
Due to restrictions, lockdowns and fear of crowded places, the COVID-19
pandemic significantly reduced the number and extent of people's
movements. Existing research suggests that the impact of the pandemic on
mobility was unequal across locations and socioeconomic groups. However,
evidence and analysis of long-term changes in mobility is scarce,
especially in less developed countries. In this study we use location
data from Meta-Facebook users to analyse patterns of urban mobility in X
Functional Urban Areas from South American countries. Findings reveal a
general decrease in mobility during the pandemic, with gradual recovery
trends up to April 2023. However, socioeconomic disparities persist
through the period of analysis, with deprived areas experiencing smaller
initial losses in mobility. Inequalities in mobility levels across
socioeconomic groups generally diminish over time, although they persist
in all the countries included in the analysis. This research highlights
the importance of understanding the long-term effects of the pandemic on
mobility and the need for equitable policy responses to address societal
challenges in urban areas.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, frame hidden, breakable, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

Spatial human mobility is key to creating sustainable, livable and
inclusive cities. At the societal level, spatial mobility enables the
transfer of knowledge, skills and labour to places they are needed.
Spatial mobility also shapes service and transport demand across urban
spaces, and enables the monitoring and control of transmissible
diseases. At the individual level, mobility enables people to access and
achieve opportunities and aspirations in space. Understanding spatial
human mobility is thus important to supporting appropriate policy
responses to address societal challenges relating to carbon emissions,
urban planning, service delivery, public health, disaster management and
transport.

The COVID-19 pandemic resulted in a notable decrease in mobility,
particularly in cities. This decrease in the overall levels of mobility
was prompted by nonpharmaceutical interventions to contain the spread of
COVID‐19, coupled with general fears of visiting crowded public
spaces.Especially during lockdowns, mobility recorded reductions in the
frequency, distance and time of trips in cities across the globe. Higher
engagement with remote working, online schooling and shopping activity
reduced the need to travel for work, education, shopping and leisure,
hence giving rise to more geographically localised mobility patterns.

However, reductions in mobility levels were highly unequal reflecting
existing socioeconomic inequalities. In most countries, affluent
individuals tended to record the greatest drops in mobility levels as
they are predominantly employed in knowledge-intensive jobs which can be
done fully or partly remotely. During the COVID-19 pandemic, the
adoption of remote work reduced the need of commuting for
knowledge-intensive, non-public facing jobs. At the same time,
individuals from less privileged socioeconomic backgrounds displayed
less pronounced declines mirroring the nature of their work requiring
public-facing, face-to-face interaction, and thus a requirement for
daily work commutes.

Thus, while a growing body of empirical evidence has contributed to
advancing our understanding of the impacts of the COVID-19 pandemic on
spatial mobility within cities, existing research has focused on more
developed countries and the immediate effects of the pandemic during
2020. Less is known about the longer term patterns of resilience in
urban mobility in less developed countries. Assessing the extent to
which the intensity of movement has returned back to pre-pandemic levels
across socioeconomic groups is important to understand the potentially
unequal impacts of hybrid working in the society.

A key barrier to monitor changes in geographic mobility patterns in less
developed countries during and post the COVID-19 pandemic has been the
lack of suitable data. Traditionally census and survey data have been
employed to analyse human mobility patterns in these countries. Yet,
these data streams are not frequently updated and suffer from slow
releases, with census data for example being collected over intervals of
ten years in most countries. Traditional data streams thus lack the
temporal granularity to analyse population movements over short-time
periods and offer an up-to-date representation of the urban mobility
system. Data resulting from social interactions on digital platforms
have emerged as an unique source of information to deliver this
representation and capture human population movement in less developed
countries at scale. Particularly location data from mobile phone
applications have become a prominent source to sense patterns of human
mobility at higher geographical and temporal resolution in real time.

Drawing on a dataset of 213 million observations from Meta-Facebook
users' mobile location data, we aim to assess socioeconomic differences
in the extent and persistence of decline in urban mobility in Argentina,
Chile, Colombia and Mexico during and after the COVID-19 pandemic from
March 2020 to March 2023. We use Meta-Facebook observed data to measure
origin-destination flows from March 2020 to May 2022, and Meta Prophet
time-series forecasting machine learning algorithm to predict
origin-destination flows from June 2022 to March 2023. We use Functional
Urban Areas (FUAs) boundaries from the Global Human Settlement Layer
(GHSL) to define urban areas, and the Global Gridded Relative
Deprivation Index (GRDI) developed by NASA's Socioeconomic Data and
Applications Centre (SEDAC) from sociodemographic and satellite data
inputs. Building on existing evidence, we hypothesised that (1) urban
mobility has recovered returning to pre-pandemic levels of movement as
nonpharmaceuthical restrictions were lifted; and (2) socioeconomic
differences in urban mobility have endured the pandemic reflecting deep
societal inequalities as knowledge-intensive businesses adopt hybrid
working.

\begin{itemize}
\tightlist
\item
  Context and importance of human mobility: public health , climate
  change\\
\item
  The pandemic resulted in major changes in human mobility - unequal
  impacts across socioeconomic groups\\
\item
  Gap 1: Little work on understanding the long-term changes in mobility
  - recovery\\
\item
  Gap 2: little work on less developed countries - where inequalities
  are more pronounced\\
\item
  Aim - focus on South American countries as case study.
\end{itemize}

\hypertarget{sec-results}{%
\section{Results}\label{sec-results}}

The evolution of the levels of movement is measured with respect to a
baseline period prior to the pandemic as described in
Section~\ref{sec-methods}. For the purposes of the analysis, we
aggregate the raw movement data temporally into months and spatially
into administrative units according to various GADM levels. The analysis
focuses on administrative areas that are within the boundaries of
functional urban areas as specified by the Global Human Settlement
Layer. For each administrative area, we compute the Relative Deprivation
Index (RDI) based on data from NASA's Socioeconomic Data and
Applications Centre (SEDAC). Figure 1 displays the administrative areas
included in the study, coloured according to the Relative Index of
Deprivation. Predictions about the evolution of the levels of movement
are made using the Prophet forecasting procedure. Further details are
provided in Section~\ref{sec-methods}.

\begin{figure}

{\centering \includegraphics{figures/maps.pdf}

}

\caption{Administrative units within functional urban areas in
Argentina, Chile, Colombia and Mexico, coloured according to the average
Relative Deprivation Index.}

\end{figure}

\hypertarget{the-impact-of-covid-19-on-urban-mobility}{%
\subsection{The impact of COVID-19 on urban
mobility}\label{the-impact-of-covid-19-on-urban-mobility}}

We analyse the evolution of the relative intensity of urban mobility
with respect to a baseline period prior to the pandemic. Specifically,
movements covering a distance of at most 70 km are considered. For a
movement to be classified as urban it needs to start or end within a
functional urban area from Argentina, Chile, Colombia and Mexico. The
observed data is available for a two-year period starting in April 2020,
just after the first wave of COVID-19 pandemic cases, and ending in
March 2022. After 2022 no observations are available, however, we
generate a 12-month forecast up to March 2023 in order to gain a better
understanding of the recovery trends.

Figure 2 displays the patterns of recovery for the mobility levels in
the administrative units belonging to functional urban areas in the
countries of interest. The three lines in each panel represent the mean
levels of mobility for administrative units grouped into one of three
terciles, according to their average relative deprivation index.

\begin{figure}

{\centering \includegraphics{figures/prediction-rdi-band.pdf}

}

\caption{Patterns of recovery for urban mobility in administrative units
from Argentina, Chile, Colombia and Mexico.}

\end{figure}

Generally, there was a drop in the levels of mobility with respect to
the baseline period in all four countries. This drop was especially
large for Argentina, Chile and Colombia, with Mexico displaying a
smaller decrease in the relative number of movements with respect to the
baseline. Following the initial drop in movement, all four countries
evolve towards the recovery of baseline levels of urban mobility, with a
generally increasing trend. There are, however, fluctuations from the
general trend which manifest differently for each country. These
fluctuations mirror each other in the case of Argentina and Colombia,
where urban mobility sharply bounces back closer to pre-pandemic levels
around July of 2020. Chile and Mexico display more progressive patterns
of recovery, although Chile never reaches baseline levels. These
fluctuations are unique to each country and can be attributed to local
factors such as the effects of seasonality or the different stringency
measures imposed by the national governments during the pandemic.

From Figure 2, we observe that there is a consistent tendency in how
administrative units with varying levels of deprivation were affected by
the pandemic. For all four countries, we observe that the administrative
units in the most deprived tercile are the ones that experienced the
smallest loss in levels of mobility at the beginning of the pandemic.
Differences in the levels of mobility across relative deprivation
terciles diminish with time. Argentina and Chile stand out as the
countries with the largest differences in mobility levels for different
relative deprivation terciles.

\hypertarget{socioeconomic-deprivation-and-recovery-of-urban-mobility}{%
\subsection{Socioeconomic deprivation and recovery of urban
mobility}\label{socioeconomic-deprivation-and-recovery-of-urban-mobility}}

In this section we explore further the role of socioeconomic deprivation
in the evolution of the levels of urban mobility. For a given point in
time (i.e.~a month), we start by considering the relationship between
the number of movements relative to the pre-pandemic baseline period and
the average relative deprivation index, at the administrative unit
level. We assume that this relationship is linear and we use a linear
regression to estimate the slope and intercept characterising the line
of best fit. This is shown for April 2020 and March 2022 in the
right-hand side panels of Figure 3. After obtaining the slope and
intercept for every month, we are able to plot the evolution of these
parameters for both the observed and forecasted data, as displayed on
the left-hand-side panels of the same Figure.

\begin{figure}

{\centering \includegraphics{figures/regression-evo-nobackground.pdf}

}

\caption{Evolution of the parameters characterising the relationship
between relative deprivation index and number of movements relative to
the pre-pandemic baseline period.}

\end{figure}

We find patterns in the evolution of the estimated parameters that
characterise the relationship between the levels of urban mobility and
RDI. In Argentina, Colombia and Mexico, we observe that the slope of
this relationship evolves to become smaller over time. The tendency is
not apparent in Chile, where the slope of the relationship remains
approximately the same despite the temporary fluctuations. The slope
captures the extent of differences in the level of urban mobility across
administrative units with varying levels of socioeconomic deprivation.
It can therefore be regarded as a measure of inequality in mobility
patterns across socioeconomic groups. A slope equal to zero would mean
that all administrative units display the same intensity of movement
regardless of their socioeconomic deprivation levels. Given the patterns
observed in Argentina, Colombia and Mexico, we find that at the
beginning of the pandemic there were notable inequalities between
socioeconomic groups in terms of the levels of urban mobility. While it
has taken more than two years for Argentina and Mexico to close the gap
(their slope is close to zero from spring 2022), inequalities persist in
Chile and Colombia as of March 2023.

The intercept of the relationship displays stronger patterns, which are
consistent across the four countries. The intercept estimates the urban
mobility levels that would be observed in administrative units where
relative deprivation is zero. The intercept was below the baseline level
at the early stages of the pandemic. As observed in Figure 3, while
there are some differences between countries in the evolution of the
intercept, the general tendency is for the intercept to increase. While
Argentina and Mexico reach values that are closer to the baseline
towards the end of the forecast period, the intercept for Chile and
Colombia remains lower. Therefore, if there were areas with no
socioeconomic deprivation, we would have seen a recovery in the levels
of mobility, especially in Argentina and Mexico

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

Using location data from Meta-Facebook users, our study aimed to examine
the evolution of patterns of mobility across socioeconomic groups in
functional urban areas from Argentina, Chile, Colombia and Mexico from
April 2020 to March 2023, following the COVID-19 pandemic. We found a
systematic decline in population movement levels during the
implementation of non-pharmaceutical interventions in April 2020, with
the largest reductions observed in the most affluent administrative
units within functional urban areas from Argentina, Chile and Colombia.
While mobility rebounded closer to pre-pandemic levels approximately two
years later after COVID-19 restrictions eased, movement intensity
remained below pre-pandemic levels in Chile and generally in areas where
the relative deprivation index is lower. Furthermore, we found that at
the beginning of the pandemic there were inequalities between
socioeconomic groups in terms of the levels of urban mobility. While it
has taken more than two years for Argentina and Mexico to gradually
close the gap, inequalities persist in Chile and Colombia as of March
2023 according to our estimated data.

We focused the analysis on short-distance movements in urban areas,
which are typically considered to capture local and routine mobility
(Owen and Green 1992). However, due to the characteristics of the
dataset, we are unable to distinguish movements of this nature from
short-distance movements that involve a permanent change of place of
residence. Future work could extend and validate our work identifying
changes to moves relating permanent and temporary displacements, and
triangulating alternative DFD and traditional data sources, such as
censuses and representative surveys collecting information on
residential location. Further research is also needed to establish the
cause of the observed changes in the spatial patterns of internal
population movement since the start of the COVID-19 pandemic.
Understanding these causes can assist in anticipating long-term
structural changes in the direction of mobility flows in the coming
years.

Additionally, accessing and working with DFD to measure human mobility
and migration is challenging. DFD are known to suffer from biases and
issues of statistical representation reflecting differences in
technology usage and accessibility (Rowe 2023). While concerted efforts
to develop data services and methodological frameworks to leverage on
DFD exist in the Global North {[}REF{]}, limited progress has been done
in Latin America. National government and statistical agencies have a
prime position to promote the creation of data sharing partnerships with
private companies and create the digital infrastructure to ensure equal
and secure access to DFD sources for the social good. With academia,
these agencies could play a major role in promoting the creation of a
methodological framework for the use and analysis of DFD, setting common
definitions, metrics, and standard approaches to measure and correct
data biases and validate data analyses. Location data have been proven
key to monitor and develop appropriate policy responses to population
displacement and movement during natural disasters, wars and epidemics
(Green, Pollock, and Rowe 2021). We envision resources like those
created by the European Commission Joint Research Centre (2022) and the
UNSD (2019), identifying sources of non-traditional data and
establishing methodological guidelines for the use of mobile phone data
for official mobility statistics. Existing resources have a global
character, we call for initiatives with a regional focus accounting for
differences in the local landscape of data availability and digital
technology usage.

\hypertarget{sec-methods}{%
\section{Methods}\label{sec-methods}}

\hypertarget{meta-facebook-movement-data}{%
\subsection{Meta-Facebook movement
data}\label{meta-facebook-movement-data}}

To capture population movements, we used anonymised aggregate mobile
phone location data from Meta users for Argentina, Colombia, Chile and
Mexico, covering a 24-month period from April 2020 to March 2022. We
used the dataset Facebook Movements created by Meta and accessed through
their Data for Good Initiative (https://dataforgood.facebook.com). The
data are built from Facebook app users who have the location services
setting turned on on their smartphone. Prior to releasing the datasets,
Meta ensures privacy and anonymity by removing personal information and
applying privacy-preserving techniques (Maas et al.~2019). Small-count
dropping is one of these techniques. A data entry is removed if the
population or movement count for an area is lower than 10. The removal
of small counts may mean that population counts in small sparsely
populated areas are not captured. A second technique consists in adding
a small undisclosed amount of random noise to ensure that it is not
possible to ascertain precise, true counts for sparsely populated
locations. Third, spatial smoothing using inverse distance-weighted
averaging is also applied is applied to produce a smooth population
count surface. The Facebook Movements dataset offers information on the
total number of Facebook users moving between and within spatial units
in the form of origin-destination matrices. The data is temporally
aggregated into three daily 8-hour time windows (i.e.~00:00-08:00,
08:00-16:00 and 16:00-00:00). The dataset includes a baseline capturing
the number of movements before COVID-19 based on a 45-day period ending
on March 10th 2020. The baseline is computed using an average for the
same time of the day and day of the week in the period preceding March
10th. For instance, the baseline for Monday 00:00-08:00 time window is
obtained by averaging over data collected on Mondays from 00:00 to 8:00
for the 45-day period. Details about the baseline can be found in Maas
et al.~(2019). The Bing Maps Tile System developed by Microsoft
(Microsoft) is used a spatial framework to organise the data. The Tile
System is a geospatial indexing system that partitions the world into
tile cells in a hierarchical way, comprising 23 different levels of
detail (Microsoft). At the lowest level of detail (Level 1), the world
is divided into four tiles with a coarse spatial resolution. At each
successive level, the resolution increases by a factor of two. The data
that we used are spatially aggregated into Bing tile levels 13. That is
about 4.9 x 4.9km at the Equator (Maas et al.~2019).

\hypertarget{spatiotemporal-data-aggregation}{%
\subsection{Spatiotemporal data
aggregation}\label{spatiotemporal-data-aggregation}}

Since the focus of this work is on urban mobility, we focus the analysis
on Functional Urban Areas (FUAs), defined by the Global Human Settlement
Layer. The spatial extent of the FUAs is often large and may include
several towns and neighbourhoods displaying a variety of socioeconomic
characteristics, hence using FUAs as the spatial units of aggregation
would considerably mask the heterogeneity in the mobility patterns.
While the original unit of aggregation for the movement data, i.e.~the
tiles from the Bing Maps Tile System, offers the highest degree of
spatial granularity available, the interpretation of findings from an
analysis based on administrative units is often more valuable. For this
reason, we perform a spatial join to aggregate the movement data into
administrative units at the GADM level 2 or 3. Only flows of people
starting or ending within the boundaries of a FUA are considered.

While the original movement data is originally aggregated into 8-hour
windows, this resolution is too fine for our analysis. Since the
analysis is focused on the longer-term evolution of patterns of urban
mobility, we aggregate the movement data by month.

In our analysis, we use the Relative Deprivation Index (RDI) as a
measure of socioeocnomic deprivation. The RDI data is made available via
NASA's Socioeconomic Data and Applications Centre (SEDAC), with a
spatial resolution of 1km pixels. We perform a spatial join of the
gridded data and the administrative units and compute the average RDI
within each of these units.

\hypertarget{time-series-analysis}{%
\subsection{Time series analysis}\label{time-series-analysis}}

For the countries included in the analysis, the movement data is
available until March 2022. In order to gain a better understanding of
the recovery trends after the pandemic, we generate a 12-month forecast
up to March 2023. This is done using Prophet, a procedure to forecast
time series data based on an additive model where non-linear trends can
be fit with seasonality effects, holiday effects and other external
factors such as the stringency index. For the analysis, only seasonality
effects are included as they yield the most realistic predictions.

In Figures X and Y, outliers are removed. Outliers are defined as
administrative units which, at any given month, have a z-score greater
than 4 for the relative intensity of movement. Furthermore, a
Savitzky-Golay filter is applied to smooth the time series data, using a
length of 4 units for the filter window and order 2 polynomials to fit
the samples.

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}