### 2024-10-23 manuscript/manuscript/supplementary.qmd
manuscript-section-results-i

---
title: "**Supplementary material**"
author:
  - name: Carmen Cabrera-Arnau
    affil-id: 1
  - name: Francisco Rowe
    affil-id: 1
  - name: Miguel González-Leonardo
    affil-id: 2
  - name: Andrea Nasuto
    affil-id: 1
  - name: Ruth Neville
    affil-id: 1
affiliations: 
  - id: 1
    name: Geographic Data Science Lab, Department of Geography and Planning, University of Liverpool, Liverpool, UK
  - id: 2
    name: Centre for Demographic Urban and Environmental Studies, El Colegio de México, Ciudad de México, México

format: 
  pdf:
    fontsize: 11pt
    geometry:
      - left=2.54cm
      - right=2.54cm
    header-includes:
      \usepackage{lipsum}
      \usepackage{libertine}
    keep-tex: true
    template-partials: 
      - title.tex
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
        \usepackage{multirow}
        \usepackage{makecell}
number-sections: true
editor: visual
csl: nature.csl
bibliography: refs.bib
editor_options:
  markdown:
    wrap: sentence
---

<!--# FR: A few questions to think about: Have we validated the data? i.e. tested their extent of representativeness? Shall we do something quick using census data for each country comparing overall populations?  -->

# Proportion of missing records across population density and socioeconomic deprivation classes,

## Facebook Population baseline data

```{=tex}
\begin{table}[ht]
 \centering 
 \captionsetup{width=\textwidth}
 \caption{Percentage of missing Facebook Population data records during the baseline period, across population density and socioeconomic deprivation classes (class 1 is the most sparsely populated and the most deprived).}

\begin{tabular}{cccccc}
  \multirow{2}{*}{\textbf{Country}} & \multirow{2}{*}{\textbf{Records count}} & \multirow{2}{*}{\textbf{Variable}} & \multirow{2}{*}{\textbf{Class 1}} & \multirow{2}{*}{\textbf{Class 2}} & \multirow{2}{*}{\textbf{Class 3}}\\
  & & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Argentina \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Chile \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Colombia \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Mexico \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\

 \end{tabular}
\vspace{5mm}
\end{table}
```
## For Facebook Movements baseline data

```{=tex}
\begin{table}[ht]
 \centering 
 \captionsetup{width=\textwidth}
 \caption{Percentage of missing Facebook Movement data records during the baseline period, across population density and socioeconomic deprivation classes of origin tile (class 1 is the most sparsely populated and the most deprived).}

\begin{tabular}{cccccc}
  \multirow{2}{*}{\textbf{Country}} & \multirow{2}{*}{\textbf{Records count}} & \multirow{2}{*}{\textbf{Variable}} & \multirow{2}{*}{\textbf{Class 1}} & \multirow{2}{*}{\textbf{Class 2}} & \multirow{2}{*}{\textbf{Class 3}}\\
  & & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Argentina \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Chile \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Colombia \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Mexico \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\

 \end{tabular}
\vspace{5mm}
\end{table}
```
## For Facebook Movements during crisis

```{=tex}

\begin{table}[ht]
 \centering 
 \captionsetup{width=\textwidth}
 \caption{Percentage of missing Facebook Movement data records during the crisis period, across population density and socioeconomic deprivation classes of origin tile (class 1 is the most sparsely populated and the most deprived).}

\begin{tabular}{cccccc}
  \multirow{2}{*}{\textbf{Country}} & \multirow{2}{*}{\textbf{Records count}} & \multirow{2}{*}{\textbf{Variable}} & \multirow{2}{*}{\textbf{Class 1}} & \multirow{2}{*}{\textbf{Class 2}} & \multirow{2}{*}{\textbf{Class 3}}\\
  & & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Argentina \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Chile \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Colombia \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \Xhline{1.5pt}

  \multirow{5}{*}{ \;\; Mexico \;\; } & \;\;Total\;\; & & & & \\ 
  \cline{2-6}
  & \multirow{2}{*}{Missing} & \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\
  \cline{2-6}
  & \multirow{2}{*}{Percentage missing} &  \;\;Pop. density\;\; & & & \\
  & & \;\;Socioec. deprivation\;\; & & & \\

 \end{tabular}
\vspace{5mm}
\end{table}
```
# References