### 2024-10-27 manuscript/manuscript-nature/manuscript-nature.qmd
manuscript-figures

---
title: Data Descriptor for Nature Scientific Data
author: 
## listed alphabetically
    - name: Carmen Cabrera-Arnau
      affiliation: 
        name: "1,* "
    - name: Francisco Rowe
      affiliation: 
        name: "1"
    - name: Miguel González-Leonardo
      affiliation: 
        name: "2"
    - name: Andrea Nasuto
      affiliation: 
        name: "1"
    - name: Ruth Neville
      affiliation: 
        name: "1"
authornotes:
    - note: "1"
      text: Geographic Data Science Lab, Department of Geography and Planning, University of Liverpool, Liverpool, UK
    - note: "2"
      text: Centre for Demographic Urban and Environmental Studies, El Colegio de México, Ciudad de México, México
    - note: "*"
      text: "`c.cabrera-arnau@liverpool.ac.uk`"
format:
  scientific-data-pdf: default
date: last-modified
abstract: |
  COVID-19 triggered a reduction in the frequency and extent of people's movement. Existing evidence suggests that while the impact of the pandemic on mobility was widespread, the extent of this impact was unequally felt across socioeconomic groups in the early stages of the pandemic. Here, we find that the most deprived locations have experienced a more accelerated recovery towards pre-pandemic levels of mobility in the long term. Furthermore, the socioeconomic disparities in the patterns of mobility triggered by the first outbreak of COVID-19, have persisted as of April 2023. These findings are based on the analysis of time-series mobility data corresponding to X urban areas from Latin American countries collected from Meta-Facebook users upon their consent. Our research highlights the importance of timely mobility data with high spatiotemporal resolution to understand the long-term effects of the pandemic and to inform equitable policy responses that address societal challenges in urban areas.

bibliography: bibliography.bib
---

# Main {.unnumbered}

<!--# CCA: Fr to rewrite section -->

Spatial human mobility is key to creating sustainable, livable and inclusive cities.
At the societal level, spatial mobility enables the transfer of knowledge, skills and labour to places they are needed [@ackers2005].
Spatial mobility also shapes service and transport demand across urban spaces [@chen2016], and enables the monitoring and control of transmissible diseases [@belik2011].
At the individual level, mobility enables people to access and achieve opportunities and aspirations in space [@klugman2009human].
Understanding spatial human mobility is thus important to supporting appropriate policy responses to address societal challenges relating to carbon emissions, urban planning, service delivery, public health, disaster management and transport [@barbosa2018; @chinazzi2020].

The COVID-19 pandemic resulted in a notable decrease in mobility, particularly in cities [@nouvellet2021].
Coupled with fears of contagion in crowded public spaces, nonpharmaceutical interventions to contain the spread of COVID‐19 prompted this decrease in the overall levels of mobility [@nouvellet2021; @rowe2023].
Especially during lockdowns, mobility recorded reductions in the frequency, distance and time of trips in cities across the globe [@abdullah2020; @bonaccorsi2020; @abu-rayash2020; @lee2023].
Higher engagement with remote working, online schooling and shopping activity reduced the need to travel for work, education, shopping and leisure, hence giving rise to more geographically localised mobility patterns [@borkowski2021].

However, reductions in mobility levels were highly unequal reflecting existing socioeconomic inequalities in our societies [@chang2020].
In most countries, affluent individuals tended to record the greatest drops in mobility levels as they are predominantly employed in knowledge-intensive jobs which can be done fully or partly remotely [@fraiberger2020; @bonaccorsi2020; @weill2020; @dueñas2021; @santana2023].
During the COVID-19 pandemic, the adoption of remote work reduced the need of commuting for knowledge-intensive, non-public facing jobs [@florida2021].At the same time, individuals from less privileged socioeconomic backgrounds displayed less pronounced declines mirroring the nature of their work requiring public-facing, face-to-face interaction, and thus a requirement for daily work commutes [@dueñas2021; @santana2023].

Thus, while a growing body of empirical evidence has contributed to advancing our understanding of the impacts of the COVID-19 pandemic on spatial mobility within cities, existing research has focused on more developed countries and the immediate effects of the pandemic during 2020.
Less is known about the longer term patterns of resilience in urban mobility in less developed countries extending beyond this period [@rowe2023a].
Urban spaces have changed considerably since then, from going through waves of high COVID-19 fatality, infections, school and business closures to the removal of all COVID-19 restrictions as the UN World Health Organization (WHO) declared an end to the pandemic as a public health emergency; yet, different configurations of hybrid working have remained in the norm across many sector of the economy [@barrero2021; @aksoy2022].
Thus, assessing the extent to which the level of mobility has returned back to the pre-pandemic baseline level across socioeconomic groups is important to understand the potentially unequal long-term impacts of hybrid working.

A key barrier to monitor changes in geographic mobility patterns in less developed countries during and post the COVID-19 pandemic has been the lack of suitable data [@rowe2023a].
Traditionally census and survey data have been employed to analyse human mobility patterns in these countries [@green2021].
Yet, these data streams are not frequently updated and suffer from slow releases, with census data for example being collected over intervals of ten years in most countries [@bell2014].
Traditional data streams thus lack the temporal granularity to analyse population movements over short-time periods and to offer an up-to-date representation of the urban mobility system [@rowe2023big].
Data resulting from social interactions on digital platforms have emerged as an unique source of information to deliver this representation and capture human population movement in less developed countries at scale [@rowe2023big].
Particularly location data from mobile phone applications have become a prominent source to sense patterns of human mobility at higher geographical and temporal resolution in real time [@calafiore2023].

Drawing on a dataset of 213 million observations from Meta-Facebook users' mobile location data, we aim to assess socioeconomic differences in the extent and persistence of decline in urban mobility in Argentina, Chile, Colombia and Mexico during and after the COVID-19 pandemic from March 2020 to March 2023.
We use Meta-Facebook data to measure origin-destination flows from March 2020 to May 2022, and Meta Prophet time-series forecasting machine learning algorithm [@prophet17] to predict origin-destination flows from June 2022 to March 2023.
<!--# Perhaps say how predictions have been validated --> We use Functional Urban Areas (FUAs) boundaries from the Global Human Settlement Layer (GHSL), developed by the European Comission's Joint Research Centre [@GHSL19] to define urban areas; and the Global Gridded Relative Deprivation Index (GRDI) developed by NASA's Socioeconomic Data and Applications Centre [@RDI2022] from sociodemographic and satellite data inputs.
Building on existing evidence [e.g. @roweCalafiore2022; @wang2022], we hypothesised that (1) urban mobility has recovered returning to the pre-pandemic baseline level of movement as nonpharmaceuthical restrictions were lifted; and, that (2) socioeconomic differences in urban mobility have endured the pandemic reflecting deep societal inequalities as knowledge-intensive businesses adopt hybrid working.
<!--# Once we have a more definitive version of the results, highlight our novel contributions here. -->

Latin America provides an ideal test-bed for testing these hypotheses because of its exceptionally high levels of inequalities [@de2004inequality; @carranza2023wealth] and urbanisation [@united2023world].
Half of the 20 most unequal countries in the planet are in this region.
The average income Gini index of the region is 4 percentage points higher than that of Africa and 11 higher than China [@milanovic2016global], and cities display some of the starkest inequalities [@habitat2022world].
Currently, over 80% of the population in Latin America live in urban areas.
By 2050, this share is predicted to reach 89%, with the largest share concentrating in a few megacities [@habitat2022world].
Developing an understanding of human mobility in Latin America is thus important to support sustainable and inclusive spaces [@habitat2022world].

# Results

## Analysing the variability in the impact of COVID-19 on mobility across population groups

We focus on capturing how mobility patterns for different population groups have been impacted by COVID-19 in Latin America.
We consider populations groups by classifying the spatial units of analysis into categories according to their population density and their socioeconomic level.
More details for the classification method are provided in section X.
The two criteria for classification are chosen due to their relevance for ...
We use five population density categories and three socioeconomic categories defined according to the relative deprivation index.
The geographic distribution of categories for each of the countries is displayed in Figures @Fig1-ARG a), @Fig1-CHL a) and @Fig1-COL a) for Argentina, Chile and Colombia respectively.

A key challenge in the in the analysis of population counts and movements with Facebook user data is the absence of small count records, which is a result of privacy-protection techniques applied to ensure that the location of individuals or small groups cannot be identified.
As a consequence, the countries of the analysis have locations with null values for Facebook population counts and flows.
These missing data are not distributed randomly.
For example, the missing values for FB population counts display high spatial autocorrelation as shown in Supplementary Figure SF1.
Therefore, simply removing the missing records from the analysis could lead to geographically biased results (Afghari et al., 2019).
To address this, we designed a data processing method to estimate the missing values, both for Facebook population counts and flows.
Furthermore, this method also applies a correction factor to eliminate the fluctuations in the daily number of observations, assuming that the representativeness of Facebook data across spatial units remains stable during the study period.
The data processing method is described in section X.

Even after applying the imputation method to estimate missing values, we acknowledge that Facebook data may still overrepresent certain population groups while underrepresenting others.
For example, Figure @Fig1 shows, for each country, the proportion of the population across the various analysis categories.
These comparisons use WorldPop population estimates and Facebook population counts, with the latter reflecting active users on an average day during a pre-pandemic baseline period (see Section X of the methodology for details).
In all three countries, notable discrepancies appear between the population distributions according to WorldPop and Facebook data.
For instance, in Argentina and Chile, WorldPop data indicates a higher proportion of people living in low-density areas, suggesting that Facebook data underrepresents populations in these regions.
Similarly, in these countries, Facebook data shows an overrepresentation of the most affluent socioeconomic group.
While addressing this kind of representativity bias is beyond the scope of this paper, we recognise its significance and the importance of addressing it in future analyses.
This issue is revisited in the Discussion section.

![Classification of spatial units into categories by population density and by relative deprivation index. Spatial distribution of categories and population share in each category, by country.](figures/Fig1.png){#Fig1}

## COVID-19’s impact on human mobility displays consistently uneven patterns across population groups in Latin American countries

The analysis of mobility is done using Facebook movement data that has been pre-processed according to our methodology.
Figure @Fig2 shows changes in the intensity of movement measured as the percentage change in the number of inflows relative to pre-pandemic baseline levels, at two points during the period of analysis, May 2020 and March 2022.
These changes show remarkably consistent patterns across countries for both types of classification.

Focusing first on the early pandemic days, we observe a decrease in the number of inflows for all countries and most categories of analysis.
This decrease displays a systematic gradient across categories of analysis whereby the more densely populated or the more affluent an area is, the higher the losses in the number of inflows.
Differences in the levels of mobility across population density or RDI categories tend to diminish with time, with a less pronounced gradient across categories of analysis approximately two years later.

These patterns can be attributed to...
remote work, movement to rural areas, affluent people being more able to abide to lockdowns, etc...
and the patterns are consistent with what has been found in ...

![Changes in the intensity of inflows as the percentage of change with respect to the pre-pandemic baseline levels, for each country and category of analysis.](figures/Fig2.png)

## Recovery trajectories towards baseline mobility levels are heterogeneous across population categories in Latin America

In this section we explore trends in the intensity of movements over time across population groups.
Generally, all categories of analysis experienced an initial drop in the number of movements followed by an upwards general trend of recovery towards baseline levels of mobility.
The extent of the initial drop and the speed of recovery varied by country and by analysis category.
These general patterns can be visually explored in Figure @Fig3 and @Fig4 for a classification by population density category and by RDI respectively.
Below we further explore these patterns by following a three-step method consisting in i) time series decomposition into general trend, seasonality and noise, ii) modelling the general trend , iii) analysing the factors that influence the values of the parameters estimated in ii) to characterise the general trend.

![](figures/Fig3.png)

![](figures/Fig4.png)

# Discussion {#sec-discussion}

Using location data from Meta-Facebook users, our study aimed to examine the evolution of patterns of mobility across socioeconomic groups in functional urban areas from Argentina, Chile, Colombia and Mexico from April 2020 to March 2023, following the COVID-19 pandemic.
We found a systematic drop in the number of population movements in April 2020, with the largest reductions observed in the most affluent administrative units within functional urban areas (FUAs) from Argentina, Chile and Colombia.
While mobility rebounded closer to pre-pandemic levels approximately two years later, when COVID-19 restrictions eased, the number of movements remained below pre-pandemic in Chile.
Furthermore, we found that at the beginning of the pandemic there were inequalities between socioeconomic groups in terms of the levels of urban mobility.
While it has taken more than two years for Argentina and Mexico to gradually reduce gap, inequalities persist as of March 2023, especially in Chile and Colombia according to our estimated data.

We focused the analysis on short-distance movements in urban areas, especifically those covering 70 km or less.
These journeys are typically considered to represent local and routine mobility [@owen1992migration] .
However, due to the characteristics of the Meta-Facebook movement data, we are unable to distinguish the purpose of these short-distance movements.
Hence, some of our data could be capturing journeys that involve a permanent change of place of residence.
Our work therefore motivates the need to answer questions regarding the validity of digital footprint data for the analysis of human mobility.
Further research should focus on inferring more specific information about the nature of the journeys, following similar approaches to those proposed by [@LaurenAlexander15; @MengChuishi17; @NiluferSariAslam21 or @Cabrera-Arnau2023], and quantifying the extent to which the digital footprint data mirrors the true mobility patterns.

Conducting research on urban mobility using digital footprint data is not straightforward, due to the challenges in accessing and working with unstructured data sets which are often subject to biases and statistical representation issues.
These biases often arise from inequalities in access and usage of digital technologies across demographic groups [@RoweBigData23].
Despite these challenges, the data and analysis that we used for this work provide evidence for non-trivial patterns that are consistent across four countries in Latin America and with other parts of the world.
Our findings highlight the dynamic interplay between socioeconomic status and urban mobility, and shall be used to motivate and inform the public debate regarding the deep societal consequences of urban mobility disparities on the wider socioeconomic landscape of Latin American countries.

In conclusion, we argue that this work goes beyond the analysis of specific patterns by demonstrating the potential of digital footprint data for policy-relevant research on human mobility at an unprecedented level of spatiotemporal granularity.
While we have seen a rise in initiatives to improve data services and methodological frameworks to facilitate the use of digital footprint data for social good, progress is still limited, especially in some parts of the world including Latin America.
It is in the hands of governments and public organisations to prioritise the maximisation the societal benefits that digital footprint data has to offer.
This includes engaging in activities such as building strategic partnerships with commercial data-holders and academic institutions to establish a unified framework for the use of digital footprint data in policy and research.
In particular, we call for the creation of resources like those developed by the European Commission Joint Research Centre [@ECJRC22] and the UN Statistics Division [@UNSD], which identify sources of non-traditional data and set methodological protocols for incorporating mobile phone data into official mobility statistics.
While current resources tend to have a global reach, we advocate for more tailored local initiatives that acknowledge disparities in regional data availability and adoption of digital technologies.

# Data and methods {.unnumbered}

## Meta-Facebook data

<!-- Include here population data too !-->

To capture population movements, we used anonymised aggregate mobile phone location data from Meta users for Argentina, Colombia, Chile and Mexico, covering a 24-month period from April 2020 to March 2022.
We used the dataset Facebook Movements created by Meta and accessed through their Data for Good Initiative (<https://dataforgood.facebook.com>).
The data are built from Facebook app users who have the location services setting turned on on their smartphone.
Prior to releasing the datasets, Meta ensures privacy and anonymity by removing personal information and applying privacy-preserving techniques [@Maas19].
Small-count dropping is one of these techniques.
A data entry is removed if the population or movement count for an area is lower than 10.
The removal of small counts may mean that population counts in small sparsely populated areas are not captured.
A second technique consists in adding a small undisclosed amount of random noise to ensure that it is not possible to ascertain precise, true counts for sparsely populated locations.
Third, spatial smoothing using inverse distance-weighted averaging is also applied is applied to produce a smooth population count surface.
The Facebook Movements dataset offers information on the total number of Facebook users moving between and within spatial units in the form of origin-destination matrices.
The data is temporally aggregated into three daily 8-hour time windows (i.e. 00:00-08:00, 08:00-16:00 and 16:00-00:00).
The dataset includes a baseline capturing the number of movements before COVID-19 based on a 45-day period ending on March 10th 2020.
The baseline is computed using an average for the same time of the day and day of the week in the period preceding March 10th.
For instance, the baseline for Monday 00:00-08:00 time window is obtained by averaging over data collected on Mondays from 00:00 to 8:00 for the 45-day period.
Details about the baseline can be found in [@Maas19].
The Bing Maps Tile System developed by Microsoft (Microsoft) is used a spatial framework to organise the data.
The Tile System is a geospatial indexing system that partitions the world into tile cells in a hierarchical way, comprising 23 different levels of detail (Microsoft).
At the lowest level of detail (Level 1), the world is divided into four tiles with a coarse spatial resolution.
At each successive level, the resolution increases by a factor of two.
The data that we used are spatially aggregated into Bing tile levels 13.
That is about 4.9 x 4.9km at the Equator [@Maas19].

## WorldPpop population data

We used data from WorldPop [@tatem2017] to classify the spatial units of analysis according to their level of urbanisation, and to estimate missing baseline values in the Facebook population data.
WorldPop offers open access gridded population estimates at a resolution as small as 3 arc-seconds approximately 100m and 1km at the Equator, respectively.
WorldPop produces these gridded datasets using top-down (i.e. dissagregating administrative area counts into smaller grid cells) or bottom-up (i.e. interpolating data from counts from sample locations into grid cells) approaches.
For the purposes of this work, we use gridded population data at a resolution of 1km$^2$ in raster format.
We perform a spatial join of the Facebook spatial units (Bing tiles) with the gridded population data and compute the sum of Worldpop populations corresponding to each of the Facebook spatial units.

## Socioeconomic deprivation data

We use the Global Gridded Relative Deprivation Index (GRDI), Version 1 (GRDIv1) data set as a measure of socioeocnomic deprivation.
The GRDI data is made available via NASA's Socioeconomic Data and Applications Centre (SEDAC), at a spatial resolution of 30 arc-seconds, or 1 km$^2$ approximately.
The index quantifies the relative levels of multidimensional deprivation and poverty, where a value of 100 represents the highest level of deprivation and a value of 0 the lowest.
We perform a spatial join of the Facebook spatial units and the gridded relative deprivation data and compute the average RDI corresponding to each of the Facebook spatial units.

## Classification of tiles according to level of urbanisation and socioeconomic deprivation

<!-- Decide whether I'm going for natural breaks or my own splitting. The disadvantage of own splitting is that there is no common logic for all four countries !-->

## Processing Facebook data

To ensure the privacy and anonymity of the users' data, Meta removes information corresponding to data entries where the population or movement count is less than 10 for a specific time or day, retaining only information about the percentage change in the number of counts with respect to the baseline period is retained [@Maas19].
Consequently, we observe many tiles where the population or movement count for either the baseline or crisis period are blank.
<!-- Justify why we do the imputation. Spatial bias (maybe show Moran's I for missing values on a selected day? Or everyday?). Small number bias, underrespresentation of dataentries with low counts that we know for sure exist !-->

### Facebook population data

To imput Facebook population baseline values, we first identify all the baseline values that are available for each spatial unit and weekday (of the three available time windows, we only consider one per day).
We then estimate the missing baseline values for each weekday, based on a linear model for the relationship between the Worldpop population and the Facebook population counts, which are fitted using ordinary least squares.
This is illustrated in Supplementary Figure \ref{fig:imput-baseline-population}.

We then use the complete baseline Facebook population values to compute missing Facebook population counts during the crisis period.
This is possible because, as mentioned above, Meta reports the percentage change in the number of counts with respect to the baseline, even if the counts are not reported due to low value.

### Facebook movement data

The imputation of Facebook movement baseline values is done according to a spatial interaction model (see e.g. [@rowe2022a]).
We considered the population flow between an origin and a destination tile, and model this variable as a function of the Facebook population count at the origin tile on the same weekday, the Facebook population count at the destination on the same weekday and the distance between origin and destination; we also included indicator variables to capture the day of the week and the pair of population density classes corresponding to the origin and destination tiles.
Mathematically, this model can be expressed as \begin{equation}
\mu_{ijw} = \beta_0 + \beta_1pop_{iw} +\beta_2pop_{jw} + \beta_3d_{ij} + \beta_4W + \beta_5C + \varepsilon
\end{equation}

where $\mu_{ijw}$ is the expectation of the flow of people from tile $i$ to tile $j$ on the weekday $w$; $\beta_0$ is an intercept $pop_{iw}$ and $pop_{jw}$ are the Facebook population counts at the origin and destination on weekday $w$ during the baseline period, $d_{ij}$ is the distance between origin and destination, $W$ is a series of indicator variables capturing the day of the week and similarly $C$ is a series of indicator variables reflecting the pair of population density classes for the origin and destination tiles, resulting in X pairs (10 classes × 10 classes minus one so it is not collinear with $\beta_0$), $\beta_{0,1,2,3,4,5}$ are model parameters to be estimated from the observed data.
The error term is denoted by \varepsilon.

To estimate the model parameters, we used a count data regression model.
Specifically, we used a negative binomial regression which is a generalised linear model (GLM) where overdispersion of the error term is assumed, i.e. the variance exceeds the mean.

We compute missing Facebook movement counts during the crisis period by considering the complete Facebook movements baseline and the percentage change in the number of counts with respect to the baseline, which is reported in the Facebook Movements datasets even when the count is not reported due to its low value.

### Correction factors??? For representativeness

## Data analysis

```{=html}
<!-- EXPLAIN - might include the things I had already written below

## Spatiotemporal data aggregation

Since the focus of this work is on urban mobility, we focus the analysis on Functional Urban Areas (FUAs), defined by the Global Human Settlement Layer (GHSL).
The spatial extent of the FUAs is often large and may include several towns and neighbourhoods displaying a variety of socioeconomic characteristics, hence using FUAs as the spatial units of aggregation would considerably mask the heterogeneity in the mobility patterns.
While the original unit of aggregation for the movement data, i.e. the tiles from the Bing Maps Tile System, offers the highest degree of spatial granularity available, the interpretation of findings from an analysis based on administrative units is often more valuable.
For this reason, we perform a spatial join to aggregate the movement data into administrative units at the GADM level 2 or 3.
Only flows of people starting or ending within the boundaries of a FUA are considered.

While the original movement data is originally aggregated into 8-hour windows, this resolution is too fine for our analysis.
Since the analysis is focused on the longer-term evolution of patterns of urban mobility, we aggregate the movement data by month.

In our analysis, we use the Relative Deprivation Index (RDI) as a measure of socioeocnomic deprivation.
The RDI data is made available via NASA's Socioeconomic Data and Applications Centre (SEDAC), with a spatial resolution of 1km pixels.
We perform a spatial join of the gridded data and the administrative units and compute the average RDI within each of these administrative units.

## Time series analysis

For the countries included in the analysis, the movement data is available until March 2022.
In order to gain a better understanding of the recovery trends after the pandemic, we generate a 12-month forecast up to March 2023.
This is done using Prophet [@prophet17], a procedure developed by Meta's Core Data Science team to forecast time series data based on an additive model.
Prophet stands out from other forecasting methods due to its ability to capture recurring seasonal trends inherent in mobility data, such as fluctuations during holidays, or changes in the general trend due to the impact of particular circumstances, such as the COVID-19 pandemic.
It is particularly intuitive to use compared to other models due to its automated trend-detection capabilities, making it accessible to users with no expertise on the underlying model, while being robust to missing data and outliers.
For the analysis in this paper, yearly seasonality effects are included as they yield the most realistic predictions.

Figures 2 and 3 have been created by removing statistical outliers.
We define these as the administrative units which, at any given month, have a z-score greater than 4 for the percentage change in the number of movements.
Furthermore, a Savitzky-Golay filter is applied to smooth the time series data, using a length of 4 units for the filter window and order 2 polynomials to fit the samples.

!-->
```
# Code availability {.unnumbered}

For all studies using custom code in the generation or processing of datasets, a statement must be included under the heading "Code availability", indicating whether and how the code can be accessed, including any restrictions to access.
This section should also include information on the versions of any software used, if relevant, and any specific variables or parameters used to generate, test, or process the current dataset.

# References

::: {#refs}
:::

# Acknowledgements {.unnumbered}

The work was supported by Research England..., project no. ...
, ‘RECAST’.

# Author information {.unnumbered}

## Authors and affiliations

## Contributions

A.A.
conceived the experiment(s), A.A.
and B.A.
conducted the experiment(s), C.A.
and D.A.
analysed the results.
All authors reviewed the manuscript.

# Ethics declarations {.unnumbered}

## Competing interests

The authors declare no competing interests.

# Supplementary information {.unnumbered}

# Rights and permissions {.unnumbered}

Springer Nature or its licensor (e.g. a society or other partner) holds exclusive rights to this article under a publishing agreement with the author(s) or other rightsholder(s); author self-archiving of the accepted manuscript version of this article is solely governed by the terms of such publishing agreement and applicable law.