### 2025-06-04 code/analysis/05-mixed-models-R.qmd
updates-code-analysis

---
title: "lme-r"
format: html
editor: visual
---

## Dependencies

```{r}
# data wrangling
library(tidyverse)
# data utilities
library(timetk)
# estimating mixed effects models
library(lme4)
library(merTools)
library(glmmTMB)
library(nlme)
# correlograms
library(ggcorrplot)
library(corrplot)
# data visualisation
library(viridis)
library(ggthemes)
library(ggpubr)
library(zoo)
library(showtext)
library(patchwork)
library(ggrepel)
# display regression equation
library(equatiomatic)
# standardise input variables
  #library(arm)
  # reporting regression results
library(broom.mixed)
library(gtsummary)
library(sjPlot)
library(modelsummary)
```

## Data import

```{r}
# Set working directory
wd <- '/Users/carmen/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/research/RECAST/latin-mobility-covid-local-files'

# Set the country
country <- 'Argentina'

# Initialize variables for country_short and country_code
country_short <- ""
country_code <- ""

# Assign country_short and country_code based on the country
if (country == 'Argentina') {
    country_short <- 'ARG'
    country_code <- 'AR'
} else if (country == 'Chile') {
    country_short <- 'CHL'
    country_code <- 'CL'
} else if (country == 'Colombia') {
    country_short <- 'COL'
    country_code <- 'CO'
} 
```

```{r}
category <- "rdi"
```

```{r}
# Read the CSV file
file_path <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_new.csv')
df <- read.csv(file_path)

df$time2 <- df$time^2
```

```{r}
dates <- read.csv(paste0(wd, '/data/outputs/', country_short, '/mov-analysis/dates.csv'))

df_stringency <- read.csv(paste0(wd, '/data/inputs/covid-stringency/owid-covid-data.csv')) %>%
  filter(location == "Chile") %>% 
  filter(date %in% dates$date) %>% 
  slice(seq(1, n(), by = 7)) %>% 
  mutate(time = seq(0, n() - 1)) %>%
  select(time, stringency_index)

df <- df %>% left_join(df_stringency, by = "time") %>% rename("O" = "cat")
```

```{r}
df_stringency$stringency_index
```

## Define models

```{r}

eq1 <- y ~ time
m1 <- glmmTMB(eq1, data = df)

###

eq1s <- y ~ time + stringency_index
m1s <- glmmTMB(eq1s, data = df)

###

# eq2 <- y ~ time + O + D
# m2 <- glmmTMB(eq2, data = df)
# 
# eq2s <- y ~ time + O + D + stringency_index
# m2s <- glmmTMB(eq2s, data = df)

eq2o <- y ~ time + O
m2o <- glmmTMB(eq2o, data = df)

eq2os <- y ~ time + O + stringency_index
m2os <- glmmTMB(eq2os, data = df)

###

# eq3 <- y ~ time + (1|O) + (1|D)
# m3 <- glmmTMB(eq3, data = df)
# m3_re <- plot_model(m3, type ="re")
# 
# eq3s <- y ~ time + (1|O) + (1|D) + stringency_index
# m3s <- glmmTMB(eq3s, data = df)
# m3s_re <- plot_model(m3s, type ="re")

eq3o <- y ~ time + (1|O)
m3o <- glmmTMB(eq3o, data = df)
m3o_re <- plot_model(m3o, type ="re")

eq3os <- y ~ time + (1|O) + stringency_index
m3os <- glmmTMB(eq3os, data = df)
m3os_re <- plot_model(m3os, type ="re")

###

# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq4 <- y ~ time + (0 + time|Of) + (0 + time|Df)
# m4 <- glmmTMB(eq4, data = df)
# m4_re <- plot_model(m4, type ="re")
# 
# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq4s <- y ~ time + (0 + time|Of) + (0 + time|Df) + stringency_index
# m4s <- glmmTMB(eq4s, data = df)
# m4s_re <- plot_model(m4s, type ="re")

df$Of <- factor(df$O)
eq4o <- y ~ time + (0 + time|Of) 
m4o <- glmmTMB(eq4o, data = df)
m4o_re <- plot_model(m4o, type ="re")

df$Of <- factor(df$O)
eq4os <- y ~ time + (0 + time|Of) + stringency_index
m4os <- glmmTMB(eq4os, data = df)
m4os_re <- plot_model(m4os, type ="re")

###

# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5 <- y ~ time + (time|Of) + (time|Df)
# m5 <- glmmTMB(eq5, data = df)
# m5_re <- plot_model(m5, type ="re")
# 
# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5s <- y ~ time + (time|Of) + (time|Df) + stringency_index
# m5s <- glmmTMB(eq5s, data = df)
# m5s_re <- plot_model(m5s, type ="re")

df$Of <- factor(df$O)
eq5o <- y ~ time + (time|Of) 
m5o <- glmmTMB(eq5o, data = df)
m5o_re <- plot_model(m5o, type ="re")

df$Of <- factor(df$O)
eq5os <- y ~ time + (time|Of) + stringency_index
m5os <- glmmTMB(eq5os, data = df)
m5os_re <- plot_model(m5os, type ="re")

###

# eq6 <- y ~ time + time2 + (time|O) + (time|D)
# m6 <- glmmTMB(eq6, data = df)
# m6_re <- plot_model(m6, type ="re")
# 
# eq6s <- y ~ time + time2 + (time|O) + (time|D) + stringency_index
# m6s <- glmmTMB(eq6s, data = df)
# m6s_re <- plot_model(m6s, type ="re")

eq6o <- y ~ time + time2 + (time|O) 
m6o <- glmmTMB(eq6o, data = df)
m6o_re <- plot_model(m6o, type ="re")

eq6os <- y ~ time + time2 + (time|O) + stringency_index
m6os <- glmmTMB(eq6os, data = df)
m6os_re <- plot_model(m6os, type ="re")

###

# eq7 <- y ~ time + time2 + (time + time2|O) + (time + time2|D)
# m7 <- glmmTMB(eq7, data = df)
# m7_re <- plot_model(m7, type ="re")
# 
# eq7s <- y ~ time + time2 + (time + time2|O) + (time + time2|D) + stringency_index
# m7s <- glmmTMB(eq7s, data = df)
# m7s_re <- plot_model(m7s, type ="re")

eq7o <- y ~ time + time2 + (time + time2|O) 
m7o <- glmmTMB(eq7o, data = df)
m7o_re <- plot_model(m7o, type ="re")

eq7os <- y ~ time + time2 + (time + time2|O) + stringency_index
m7os <- glmmTMB(eq7os, data = df)
m7os_re <- plot_model(m7os, type ="re")

```

## Save model data

```{r}
models_o <- list(
  m1,
  m2o,
  m3o,
  m4o,
  m5o,
  m6o,
  m7o
  )
tidy_results_o <- map_df(models_o, tidy, .id = "model")
tidy_results_o <- map_df(models_o, tidy, .id = "model")
file_tidy_results_o <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_o_new.csv')
write.csv(tidy_results_o, file_tidy_results_o, row.names = FALSE)

glance_results_o <- map_df(models_o, glance, .id = "model")
file_glance_results_o <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_glance_results_o_new.csv')
write.csv(glance_results_o, file_glance_results_o, row.names = FALSE)

models_os <- list(
  m1s,
  m2os,
  m3os,
  m4os,
  m5os,
  m6os,
  m7os
  )
tidy_results_os <- map_df(models_os, tidy, .id = "model")
file_tidy_results_os <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_os_new.csv')
write.csv(tidy_results_os, file_tidy_results_os, row.names = FALSE)

glance_results_os <- map_df(models_os, glance, .id = "model")
file_glance_results_os <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_glance_results_os_new.csv')
write.csv(glance_results_os, file_glance_results_os, row.names = FALSE)
```

## Save random effects data

```{r}

file_m3o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m3o_re_new.csv')
write.csv(m3o_re[[1]], file_m3o_re, row.names = FALSE)

file_m3os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m3os_re_new.csv')
write.csv(m3os_re[[1]], file_m3os_re, row.names = FALSE)

###

file_m4o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m4o_re_new.csv')
write.csv(m4o_re[[1]], file_m4o_re, row.names = FALSE)

file_m4os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m4os_re_new.csv')
write.csv(m4os_re[[1]], file_m4os_re, row.names = FALSE)

###

file_m5o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m5o_re_new.csv')
write.csv(m5o_re[[1]], file_m5o_re, row.names = FALSE)

file_m5os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m5os_re_new.csv')
write.csv(m5os_re[[1]], file_m5os_re, row.names = FALSE)

###

file_m6o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m6o_re_new.csv')
write.csv(m6o_re[[1]], file_m6o_re, row.names = FALSE)

file_m6os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m6os_re_new.csv')
write.csv(m6os_re[[1]], file_m6os_re, row.names = FALSE)

###

file_m7o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m7o_re_new.csv')
write.csv(m7o_re[[1]], file_m7o_re, row.names = FALSE)

file_m7os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m7os_re_new.csv')
write.csv(m7os_re[[1]], file_m7os_re, row.names = FALSE)

```

## Plots of RE

```{r}
m3o_re
m3os_re
m4o_re
m4os_re
m5o_re
m5os_re
m6o_re
m6os_re
m7o_re
m7os_re
```