### 2025-12-01 notebooks/analysis/05b-mixed-models-R-r1.qmd
revision

---
title: "lme-r"
format: html
editor: visual
---

## Dependencies

```{r}
# data wrangling
library(tidyverse)
# data utilities
library(timetk)
# estimating mixed effects models
library(lme4)
library(merTools)
library(glmmTMB)
library(nlme)
# correlograms
library(ggcorrplot)
library(corrplot)
# data visualisation
library(viridis)
library(ggthemes)
library(ggpubr)
library(zoo)
library(showtext)
library(patchwork)
library(ggrepel)
# display regression equation
library(equatiomatic)
# standardise input variables
  #library(arm)
  # reporting regression results
library(broom.mixed)
library(gtsummary)
library(sjPlot)
library(modelsummary)
library(brms)
```

## Data import

```{r}
# Set working directory
wd <- '/your/path/to/working/directory/'
wd <- '/Users/carmen/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/research/latin-mobility-covid'

# Set the country
country <- 'Argentina'

# Initialize variables for country_short and country_code
country_short <- ""
country_code <- ""

# Assign country_short and country_code based on the country
if (country == 'Argentina') {
    country_short <- 'ARG'
    country_code <- 'AR'
} else if (country == 'Chile') {
    country_short <- 'CHL'
    country_code <- 'CL'
} else if (country == 'Colombia') {
    country_short <- 'COL'
    country_code <- 'CO'
} 
```

```{r}
category <- "density"
```

```{r}
# Read the CSV file
file_path <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_new.csv')
df <- read.csv(file_path)

df$time2 <- df$time^2
```

```{r}
dates <- read.csv(paste0(wd, '/data/outputs/', country_short, '/mov-analysis/dates.csv'))

df_stringency <- read.csv(paste0(wd, '/data/inputs/covid-stringency/owid-covid-data.csv')) %>%
  filter(location == country) %>% 
  filter(date %in% dates$date) %>% 
  slice(seq(1, n(), by = 7)) %>% 
  mutate(time = seq(0, n() - 1)) %>%
  select(time, stringency_index)

df <- df %>% left_join(df_stringency, by = "time") %>% rename("O" = "cat")

```

## Define models

```{r}

eq1 <- y ~ time
m1 <- glmmTMB(eq1, data = df)

###

eq1s <- y ~ time + stringency_index 
m1s <- glmmTMB(eq1s, data = df)

###

# eq2 <- y ~ time + O + D 
# m2 <- glmmTMB(eq2, data = df)
# 
# eq2s <- y ~ time + O + D + stringency_index 
# m2s <- glmmTMB(eq2s, data = df)

eq2o <- y ~ time + O 
m2o <- glmmTMB(eq2o, data = df)

eq2os <- y ~ time + O + stringency_index 
m2os <- glmmTMB(eq2os, data = df)

###

# eq3 <- y ~ time + (1|O) + (1|D) 
# m3 <- glmmTMB(eq3, data = df)
# m3_re <- plot_model(m3, type ="re")
# 
# eq3s <- y ~ time + (1|O) + (1|D) + stringency_index 
# m3s <- glmmTMB(eq3s, data = df)
# m3s_re <- plot_model(m3s, type ="re")

eq3o <- y ~ time + (1|O) 
m3o <- glmmTMB(eq3o, data = df)
m3o_re <- plot_model(m3o, type ="re")

eq3os <- y ~ time + (1|O) + stringency_index 
m3os <- glmmTMB(eq3os, data = df)
m3os_re <- plot_model(m3os, type ="re")

###

# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq4 <- y ~ time + (0 + time|Of) + (0 + time|Df) 
# m4 <- glmmTMB(eq4, data = df)
# m4_re <- plot_model(m4, type ="re")
# 
# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq4s <- y ~ time + (0 + time|Of) + (0 + time|Df) + stringency_index 
# m4s <- glmmTMB(eq4s, data = df)
# m4s_re <- plot_model(m4s, type ="re")

df$Of <- factor(df$O)
eq4o <- y ~ time + (0 + time|Of) 
m4o <- glmmTMB(eq4o, data = df)
m4o_re <- plot_model(m4o, type ="re")

df$Of <- factor(df$O)
eq4os <- y ~ time + (0 + time|Of) + stringency_index 
m4os <- glmmTMB(eq4os, data = df)
m4os_re <- plot_model(m4os, type ="re")

###

# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5 <- y ~ time + (time|Of) + (time|Df) + ratio
# m5 <- glmmTMB(eq5, data = df)
# m5_re <- plot_model(m5, type ="re")
# 
# df$Of <- factor(df$O)
# df$Df <- factor(df$D)
# eq5s <- y ~ time + (time|Of) + (time|Df) + stringency_index + ratio
# m5s <- glmmTMB(eq5s, data = df)
# m5s_re <- plot_model(m5s, type ="re")

df$Of <- factor(df$O)
eq5o <- y ~ time + (time|Of) 
m5o <- glmmTMB(eq5o, data = df)
m5o_re <- plot_model(m5o, type ="re")

df$Of <- factor(df$O)
eq5os <- y ~ time + (time|Of) + stringency_index 
m5os <- glmmTMB(eq5os, data = df)
m5os_re <- plot_model(m5os, type ="re")

###

# eq6 <- y ~ time + time2 + (time|O) + (time|D) + ratio
# m6 <- glmmTMB(eq6, data = df)
# m6_re <- plot_model(m6, type ="re")
# 
# eq6s <- y ~ time + time2 + (time|O) + (time|D) + stringency_index + ratio
# m6s <- glmmTMB(eq6s, data = df)
# m6s_re <- plot_model(m6s, type ="re")

eq6o <- y ~ time + time2 + (time|O) 
m6o <- glmmTMB(eq6o, data = df)
m6o_re <- plot_model(m6o, type ="re")

eq6os <- y ~ time + time2 + (time|O) + stringency_index 
m6os <- glmmTMB(eq6os, data = df)
m6os_re <- plot_model(m6os, type ="re")

###

# eq7 <- y ~ time + time2 + (time + time2|O) + (time + time2|D) + ratio
# m7 <- glmmTMB(eq7, data = df)
# m7_re <- plot_model(m7, type ="re")
# 
# eq7s <- y ~ time + time2 + (time + time2|O) + (time + time2|D) + stringency_index + ratio
# m7s <- glmmTMB(eq7s, data = df)
# m7s_re <- plot_model(m7s, type ="re")

eq7o <- y ~ time + time2 + (time + time2|O) 
m7o <- glmmTMB(eq7o, data = df)
m7o_re <- plot_model(m7o, type ="re")

eq7os <- y ~ time + time2 + (time + time2|O) + stringency_index 
m7os <- glmmTMB(eq7os, data = df)
m7os_re <- plot_model(m7os, type ="re")

df$O  <- factor(df$O)
eq8o <- y ~ 
    s(time, k = 8) + 
    s(time, O, bs = "fs", k = 7)
m_gam <- gam(eq8o, data = df, method = "REML")

df$O  <- factor(df$O)
eq8os <- y ~ 
    s(time, k = 8) + 
    stringency_index + 
    s(time, O, bs = "fs", k = 7)
m_gams <- gam(eq8os, data   = df, method = "REML")

```

```{r}
summary(m_gam)$s.table
```

```{r}
plot(m_gam, pages = 1, rug = FALSE, scheme = 1, shade = TRUE)
```

## Save model data

```{r}
models_o <- list(
  m1,
  m2o,
  m3o,
  m4o,
  m5o,
  m6o,
  m7o, 
  m_gam
  )
tidy_fixed <- map_df(models_o, ~ tidy(.x, parametric = TRUE), .id = "model")
tidy_smooth <- map_df(models_o, ~ tidy(.x, smooth = TRUE), .id = "model")

tidy_results_o <- bind_rows(tidy_fixed, tidy_smooth) %>%
  distinct()

file_tidy_results_o <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_o_new.csv')
write.csv(tidy_results_o, file_tidy_results_o, row.names = FALSE)

glance_results_o <- map_df(models_o, glance, .id = "model")
file_glance_results_o <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_glance_results_o_new.csv')
write.csv(glance_results_o, file_glance_results_o, row.names = FALSE)

models_os <- list(
  m1s,
  m2os,
  m3os,
  m4os,
  m5os,
  m6os,
  m7os,
  m_gams
  )
tidy_fixed <- map_df(models_os, ~ tidy(.x, parametric = TRUE), .id = "model")
tidy_smooth <- map_df(models_os, ~ tidy(.x, smooth = TRUE), .id = "model")
tidy_results_os <- bind_rows(tidy_fixed, tidy_smooth) %>%
  distinct()
file_tidy_results_os <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_tidy_results_os_new.csv')
write.csv(tidy_results_os, file_tidy_results_os, row.names = FALSE)

glance_results_os <- map_df(models_os, glance, .id = "model")
file_glance_results_os <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_outflows_param_glance_results_os_new.csv')
write.csv(glance_results_os, file_glance_results_os, row.names = FALSE)
```

## Save random effects data

```{r}

file_m3o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m3o_re_new.csv')
write.csv(m3o_re[[1]], file_m3o_re, row.names = FALSE)

file_m3os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m3os_re_new.csv')
write.csv(m3os_re[[1]], file_m3os_re, row.names = FALSE)

###

file_m4o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m4o_re_new.csv')
write.csv(m4o_re[[1]], file_m4o_re, row.names = FALSE)

file_m4os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m4os_re_new.csv')
write.csv(m4os_re[[1]], file_m4os_re, row.names = FALSE)

###

file_m5o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m5o_re_new.csv')
write.csv(m5o_re[[1]], file_m5o_re, row.names = FALSE)

file_m5os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m5os_re_new.csv')
write.csv(m5os_re[[1]], file_m5os_re, row.names = FALSE)

###

file_m6o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m6o_re_new.csv')
write.csv(m6o_re[[1]], file_m6o_re, row.names = FALSE)

file_m6os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m6os_re_new.csv')
write.csv(m6os_re[[1]], file_m6os_re, row.names = FALSE)

###

file_m7o_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m7o_re_new.csv')
write.csv(m7o_re[[1]], file_m7o_re, row.names = FALSE)

file_m7os_re <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/data_trend_param_quad_m7os_re_new.csv')
write.csv(m7os_re[[1]], file_m7os_re, row.names = FALSE)
```

## Plots of RE

```{r}
# m3o_re
# m3os_re
# m4o_re
# m4os_re
# m5o_re
# m5os_re
# m6o_re
# m6os_re
# m7o_re
# m7os_re
```

## Compute recovery times for spline model. The 95% confidence intervals are computed via bootstrap.

```{r}
stringency <- FALSE
```

```{r}
# Create new time sequence covering the range of your original data
new_time <- expand.grid(
  time = seq(min(df$time), max(df$time)*100, length.out = max(df$time)*100-min(df$time)+1),
  O = levels(df$O)   # one dataset per group
)

new_time <- new_time %>%
  left_join(df[, c("time", "O", "stringency_index")], by = c("time", "O"))

new_time$stringency_index[is.na(new_time$stringency_index)] <- 0

if (stringency == FALSE) {
  new_time <- expand.grid(
  time = seq(min(df$time), max(df$time)*100, length.out = max(df$time)*100-min(df$time)+1),
  O = levels(df$O)   # one dataset per group
)
  pred_base <- predict(m_gam, newdata = new_time, se.fit = TRUE)
} else {
  pred_base <- predict(m_gams, newdata = new_time, se.fit = TRUE)
}

new_time$fit <- pred_base$fit
new_time$se  <- pred_base$se.fit

# Find FIRST root where preds(t) == target
get_root_threshold <- function(time, preds, target) {
  diff_sign <- diff(sign(preds - target))
  crossings <- which(diff_sign != 0)
  if (length(crossings) == 0) return(NA)

  i <- crossings[1]  # first crossing
  out <- tryCatch(
    uniroot(function(t) {
      approx(time, preds - target, xout = t)$y
    },
    lower = time[i],
    upper = time[i + 1])$root,
    error = function(e) NA
  )
  return(out)
}

alphas <- seq(0, 1, by = 0.05)   # adjust resolution as needed

# Bootstrap confidence intervals
set.seed(123)
n_boot <- 200   # increase to 500–1000 for publication

results_alpha <- map_df(levels(df$O), function(g) {

  group_data <- new_time %>% filter(O == g)

  # group-specific y(t=0): use earliest time in grid for that group
  idx0 <- which.min(group_data$time)
  y0_group <- group_data$fit[idx0]

  # For each alpha / target, bootstrap
  map_df(seq_along(alphas), function(j) {
    alpha   <- alphas[j]
    target <- (1 - alpha) * y0_group

    boot_roots <- replicate(n_boot, {
      # simulate one curve from uncertainty
      y_sim <- rnorm(nrow(group_data), mean = group_data$fit, sd = group_data$se)
      get_root_threshold(group_data$time, y_sim, target)
    })

    data.frame(
      O        = g,
      alpha    = alpha,
      y_target = target,
      t_alpha  = median(boot_roots, na.rm = TRUE),
      CI_lower = quantile(boot_roots, 0.025, na.rm = TRUE),
      CI_upper = quantile(boot_roots, 0.975, na.rm = TRUE),
      SD_t_alpha = sd(boot_roots, na.rm = TRUE)
    )
  })
})

results_alpha <- as_tibble(results_alpha)

```

## Save recovery times for splines model

```{r}

if (stringency == FALSE) {
  file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-gam-r1.csv')
} else {
  file_results_alpha <- paste0(wd, '/data/outputs/', country_short, '/mov-analysis/by-', category, '/recovery-times-gams-r1.csv')
}
write.csv(results_alpha, file_results_alpha, row.names = FALSE)

```

## Plot recovery times splines model for various alpha

```{r}
ggplot(results_alpha, aes(x = alpha, y = t_alpha, color = O, group = O)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper, fill = O),
              alpha = 0.15, color = NA) +
  labs(
    title = "Time to reach threshold level (1 - α)",
    x = expression(alpha),
    y = expression(t(alpha)),
    color = "Group (O)",
    fill = "Group (O)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11),
    plot.title = element_text(face = "bold", size = 16)
  )

```